\ProjectName{}: Data Imputation as a Query Optimization

### 2017-02-28 paper/relatedwork.tex
paper: pass over relatedwork

In~\cite{akande2015empirical}, Akande et al analyze the performance of various
multiple imputation techniques on the American Community Survey dataset. 
The computational difficulties of imputing on large base
tables are well known and can limit approaches. For example, Akande finds that
one approach (MI-GLM) is prohibitively expensive when attempting to impute on data that
includes variables with potentially large domains (ten categories in their case).
In contrast, \ProjectName{} allows users to specify a tradeoff between
information lost and runtime performance, allowing users to perform queries
directly on the entire dataset. Furthermore,
the query planner's imputation is guided by the requirements of each specific
query's operators, rather than requiring broad assumptions about query
workloads.

@@

Designers of data stream processing systems frequently confront missing values
and consider them carefully in query processing. Often, if sensor error is the
cause of missing values, values can be imputed with high confidence. In
\cite{fernandez2009inter}, feedback punctuation is used to dynamically
reconfigure the query plan as data continue to arrive for state-dependent
optimizations. In this framework, expensive imputations can be avoided as
real-time conditions change.

@@

\subsection{Forecasting and Databases}
Parisi et al\cite{parisi2011embedding} introduce the idea of incorporating time-series forecast operators into
databases, along with the necessary relational algebra extensions. Their work explores the theoretical
properties of forecast operators and generalizes them into a family of operators, distinguished by
the type of predictions returned. They highlight the use of forecasting for replacing missing values.
In their follow on work\cite{parisi2013temporal}, they identify various equivalence and containment
relationships when using forecast operators, which could be used to perform query plan transformations that guarantee the same result. They
explore forecast-first and forecast-last plans, which perform forecasting operations before and after traditional
relational operators, respectively.

In contrast to this work, \ProjectName{} is targeted at generic value imputation, not necessarily tailored to 
time-series. The optimizer is not based on equivalence transformations, nor are there guarantees of equal
results under different conditions. Instead, the optimizer allows users to pick their tradeoff between
runtime cost and information lost. The search space considered by our optimizer is broader, not just
forecast-first/forecast-last plans, but rather imputation operators can be placed anywhere in the query plan
(with some restrictions). The novelty of our contribution lies in the successful incorporation of
imputation operations in non-trivial query plans with cost-based optimization.

>>>>

In~\cite{akande2015empirical}, Akande et al analyze the performance of various
multiple imputation techniques on the American Community Survey dataset. 
The computational difficulties of imputing on large base
tables are well known and can limit approaches. For example, Akande finds that
one approach (MI-GLM) is prohibitively expensive when attempting to impute on data that
includes variables with potentially large domains (ten categories in their case).
In contrast, \ProjectName{} allows users to specify a tradeoff between
imputation quality and runtime performance, allowing users to perform queries
directly on the entire dataset. Furthermore,
the query planner's imputation is guided by the requirements of each specific
query's operators, rather than requiring broad assumptions about query
workloads.

@@

Designers of data stream processing systems frequently confront missing values
and consider them carefully in query processing. Often, if sensor error is the
cause of missing values, values can be imputed with high confidence. In
\cite{fernandez2009inter}, feedback punctuation is used to dynamically
reconfigure the query plan for state-dependent
optimizations as data continue to arrive. One of the applications of this framework is to avoid
expensive imputations as real-time conditions change. 
\ProjectName{} focuses on handling
missing values, rather than providing a general framework for planning feedback. Furthermore,
it provides the ability
to trade imputation quality for runtime performance. We prove our
algorithm and cost model produces an optimal set of plans to choose from,
given the constraints of the search space.

@@

\subsection{Forecasting and Databases}
Parisi et al\cite{parisi2011embedding} introduce the idea of incorporating time-series forecast operators into
databases, along with the necessary relational algebra extensions. Their work explores the theoretical
properties of forecast operators and generalizes them into a family of operators, distinguished by
the type of predictions returned. They highlight the use of forecasting for replacing missing values.
In their follow on work\cite{parisi2013temporal}, they identify various equivalence and containment
relationships when using forecast operators, which could be used to perform query plan transformations that guarantee the same result. They
explore forecast-first and forecast-last plans, which perform forecasting operations before and after executing all traditional
relational operators, respectively.

In contrast to this work, \ProjectName{} is targeted at generic value imputation, not necessarily tailored to 
time-series. The optimizer is not based on equivalence transformations, nor are there guarantees of equal
results under different conditions. Instead, the optimizer allows users to pick their tradeoff between
runtime cost and imputation quality. The search space considered by our optimizer is broader, not just
forecast-first/forecast-last plans, but rather imputation operators can be placed anywhere in the query plan
(with some restrictions). The novelty of our contribution lies in the successful incorporation of
imputation operations in non-trivial query plans with cost-based optimization.
